### 2024-08-16 article.tex
Share the draft

\documentclass[12pt,prb,aps]{revtex4-2}
\usepackage{amsmath,amssymb,physics}
\usepackage{hyperref} % Clickable links in the document
\usepackage{graphicx} % For figures with images
\usepackage{natbib} % bibliography
\setcitestyle{super, sort&compress, comma} % chemistry style
\usepackage{hyperref} % Clickable links in the document and their colors
\hypersetup{
    colorlinks=true, % false: boxed links; true: colored links
    linkcolor=cyan, % color of internal links
    citecolor=magenta, % color of links to bibliography
    filecolor=magenta, % color of file links
    urlcolor=cyan, % color of external links
    runcolor=cyan
}

\begin{document}

\title{Vibronic effects in the photoelectron spectrum of ozone}

% \author{Pawe{\l} W{\'o}jcik}
% \email[Electronic address: ]{pawel.wojcik@usc.edu}
% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}
% \author{Anna I. Krylov}
% \email[Electronic address: ]{krylov@usc.edu}
% \affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
% \author{John F. Stanton}
% \affiliation{Quantum Theory Project, Department of Chemistry, University of Florida, Gainesville, Florida 32611, USA}
% \affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, H-1518 Budapest 112, Hungary}

\date{\today}

\begin{abstract}Abstract\end{abstract}

% TODO: Add that in this work we use the Mulliken's notation for state
% symmetry

\maketitle

\section{Introduction}

Ozone is a C$_{2v}$ molecule with three normal modes: symmetric stretch,
symmetric bend, and asymmetric stretch. The asymmetric stretch is of a b$_2$
symmetry. The two lowest electronic states of the ozone cation are close in
energy. The lower state is fully symmetric while the higher one is of the
B$_2$ symmetry, the same symmetry as the asymmetric stretch. The close
energetic separation and the matching symmetry of the asymmetric stretch
results in significant vibronic effects appearing in the ozone photoelectron
spectrum. Figure~\ref{fig:ozone_intro} presents a graphical summary of this
information.

\begin{figure}
    \includegraphics[width = 16 cm]{./figures/ozone_intro}
    \caption{ 
        Two lowest states of the ozone cation and normal modes of ozone.
    }
    \label{fig:ozone_intro}
\end{figure}

\section{Methods}

We simulate the vibronic states of the ozone cation using the model
Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum (KDC
Hamiltonian).\cite{Cederbaum:LVC:84,KDC:81,Koppel:CIbookCh7:04} This is a
multi-state and multi-mode Hamiltonian defined in the basis of diabatic
states.  We consider the ozone cation in the basis of two quasi-diabatic
states coupled by one mode (mode number 3). The model also includes two
symmetric modes (modes number 1 and 2)
\begin{subequations}
    \begin{equation}
        H = H _0 \mathbf{1}
        +
        \begin{pmatrix}
            V ^{(1)}  & \lambda _3 Q _3\\
            \lambda _3 Q _3 & V ^{(2)}
        \end{pmatrix}
    \end{equation}
    \begin{equation}
        H _0 = 
        \frac{1}{2} \left(\sum _{i = 1}^3 
        - \omega _i \frac{\partial ^2}{\partial Q _i ^2 }\right)
        + \frac{1}{2}\omega _3 Q _3 ^2
    \end{equation}
    \begin{equation}
        V ^{(\alpha)} = 
        E ^{(\alpha)} 
        + \sum _{i,j,k,l \in \{1,2\}} 
            \kappa ^{(\alpha)} _i Q _i 
            + \kappa ^{(\alpha)} _{ij} Q _i Q _j 
            + \kappa ^{(\alpha)} _{ijk} Q _i Q _j Q _k 
            + \kappa ^{(\alpha)} _{ijkl} Q_i Q _j Q _k Q _l.
    \end{equation}
    \label{eq:KDC}
\end{subequations}
The Hamiltonian parameters are expanded around the equilibrium geometry of
ozone. $E ^{(\alpha)}$ are the vertical ionization energies calculated at that
geometry. $Q_i$ are the dimensionless normal coordinates of ozone. $\kappa$ are
the coefficients of expansion of the potential along the fully symmetric
coordinates. $\lambda$ are the linear diabatic couplings.  $\omega$ are the
harmonic frequencies of ozone.

We find the parameters that enter the KDC Hamiltonian using \emph{ab inito}
coupled-cluster (CC) methods and its equation-of-motion EOM-CC
extension.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,
Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} We use the CC
truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT)
as well as singles, doubles, triples and quadruples
(CCSDTQ).\cite{Matthews:ncc:2015}  We use the EOM-CC for ionization potential
(EOM-IP).~\cite{StantonGauss:EOMIP:99} We use the Ichino, Gauss and Stanton
definition of quasi-diabatic states based on the EOM-CC method
(EOM-CC-QD)~\cite{Stanton:EOMIPdeg:09}. In all CC and EOM-CC calculations we
leave all electrons correlated, in other words, we do not use the
frozen-core approximation.

Using CCSDT/ANO1 we optimize the geometry of ozone and compute its harmonic
frequencies and normal coordinates.~\cite{Almlof:ANO,Almlof:ANO:1988} At the
same geometry we compute the linear diabatic coupling $\lambda$ using
EOM-IP-CCSD-QD/ANO1.  We use EOM-IP-CCSDT/ANO1 on a grid to find the
expansion coefficients $\kappa$.

The vertical ionization energies are calculated using a composite method. The
base value is the complete basis set (CBS) extrapolation of the
EOM-IP-CCSDT/cc-pCVnZ energies with n = 5, 6.~\cite{Woon:95:CCBS} These values
are augmented with two corrections: the $\Delta$Q correction in the cc-pwCVTZ
basis set and the relativistic correction calculated using
EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} We introduce an error
estimates to the reported vertical energies. For the error estimate of the
extrapolated CBS value we use half of the absolute value of the difference
between the best \emph{ab initio} value and the extrapolated value. For the
remaining corrections we use as the error estimate half of the absolute value
of the correction.

To compare the simulated photoelectron spectrum to the experimental one, we
use the oscillator strengths ratio A$_1$:B$_2$ of $1$:$1.35$.\cite{KDC:O3:92}
Additionally, the stick spectrum is broadened with the Lorenzian envelopes
normalized to the peaks' intensities
\begin{equation}
    f _{envelope}(x, x _i, I _i) = 
    \frac{I_i}{(x-x _i)^2 + (\gamma/ 2) ^2}.
    \label{eq:lorentzian}
\end{equation}
$x _i$ is the position of the spectral peak, $I _i$ is its intensity and
$\gamma$ is the peak's width.

We use the \textsc{xsim} program of Dr. Stanton to simulate the spectrum using
the basis of 50 harmonic states in each mode and 6000 iterations of the
Lanczos procedure. All remaining \emph{ab initio} calculations are completed
using \textsc{CFOUR}.~\cite{cfour, cfour:2020}

\section{Results}

The optimized geometry of ozone gives the bond lengths of $1.270$~\AA{} and
the bond angle of $116.9^\circ$. The two symmetric normal modes have
frequencies $1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch
has frequency equal to $1129$~cm$^{-1}$. The computed value of the linear
diabatic coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical
ionization energy for the first excited state, $E^{(1)}$, computed using the
composite method described above is equal to $12.827$~eV. Our error estimate
for this value is $30$~meV, see Table~\ref{tab:vertical_ionization_energy} for
details.  We note that the convergence of the vertical ionization gap between
the two states is much faster and we estimate this value as equal to
$123\pm8$~meV, see Table~\ref{tab:vertical_gap} for details.

\begin{table}
    \caption{
        Vertical ionization energies with the error estimates, eV.
    }
    \label{tab:vertical_ionization_energy}
    \begin{center}
        \begin{tabular}[c]{|l|rr|r|}
            \hline
            Contribution  & $^2$A$_1$ & $^2$B$_2$ & Error estimate \\ \hline
            CBS           & 12.872    & 12.981    & 0.02 \\
            $\Delta$Q     & -0.037    & -0.021    & 0.02 \\
            Relativistic  & -0.008    & -0.010    & 0.005 \\ \hline
            Sum           & 12.827    & 12.950    & 0.03 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{table}
    \caption{
        The energy gap between vertical ionization energies of the $^2$A$_1$
        and $^2$B$_2$ states with error estimates, meV. See caption of
        Table~\ref{tab:vertical_ionization_energy}.
    }
    \label{tab:vertical_gap}
    \begin{center}
        \begin{tabular}[c]{|l|r|r|}
            \hline
            Contribution             &  Gap    & Error estimate \\ \hline
            EOM-CCSDT/CBS            &  108.8  & 0.9 \\
            $\Delta$Q/pwCVTZ         &   16.5  & 8 \\
            Relativistic/CCSD/pwCVTZ &   -2.2  & 1.1 \\ \hline
            Final value, meV         &  123    & 8 \\
            Final value, cm$^{-1}$   &  990    & 65 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{figure}
\includegraphics[width = 8 cm]{figures/sim_vs_Dyke}
\caption{
    Comparison of the experimental (solid black line) and the simulated
    spectrum. In the simulated spectra the plot is generated with peak width
    $\gamma = 30$~meV. The simulated spectrum is shifted towards higher
    energies by $21$~meV.
}
\label{fig:sim_vs_dyke}
\end{figure}

Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
experimental one taken from reference~\cite{dyke:O3:74}. A comparison of the
simulated and experimental spectra is presented on
Figure~\ref{fig:sim_vs_dyke} and shows a good match. Peak A is known to be a
hot band.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
increase in the intensities of peaks B, C, D, and E. The spacing between these
peaks is also well reproduced. Drop in the intensity of peak F is also
captured by the simulation. Starting from peak G the simulation shows
discrepancy with the experiment. A sudden drop in the intensity past the peak
H is not observed in the simulation. We discuss a likely source of this
mismatch later.

\begin{figure}
    \includegraphics[width=10 cm]{figures/spectrum_overline}
    \caption{
        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
        scale in eV. Top axis shows energy offset from the origin in
        cm$^{-1}$. Stick spectrum shows positions and intensities of all
        simulated states. Blue color corresponds to the oscillator strength
        originating from the $^2$A$_1$ basis state while the red ones
        correspond to $^2$B$_2$. Gray vertical lines with captions on top
        indicate positions of features as measured by the PFI-ZEKE
        experiment.~\cite{Willitsch:O3ZEKE:2005} 
        $D_0$ marks the dissociation threshold of O$_3^+$.
        Gray dotted line marks the energy of the minimum of the conical
        intersection (CI).
    }
    \label{fig:ozone_overlay}
\end{figure}

Our simulation allows for an additional element of analysis of the simulated
spectrum. Figure~\ref{fig:ozone_overlay} presents the same simulated spectrum
as in Figure~\ref{fig:sim_vs_dyke} with additional information on the
decomposition of the spectrum.  All lines that contribute to the spectrum are
marked individually and in two different colors. The spectrum shows that peaks
A and B are almost purely of the $^2$A$_1$ character. Starting from peak C
each the contributions from two states are equally important, it is also clear
that each peak of the low-resolution spectrum has in fact contributions from
many vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin
the density of vibronic peaks increases significantly. This value can be
compared to the minimum of the conical intersection, which our simulation
locates at $3174$~cm$^{-1}$ above the origin ($12.92$~eV). Our simulation is
incapable of accounting for dissociation of the molecule, therefore the we
expect a discrepancy with the experiment as the energy gets closer to the
dissociation threshold located at $4840$~cm$^{-1}$ above the
origin.~\cite{Willitsch:O3ZEKE:2005}

\begin{figure}
    \includegraphics[width=10cm]{figures/spectrum_assigned.pdf}
    \caption{
        Simulated photoelectron spectrum of ozone, but with no account given
        to the vibronic coupling effect.
    }
    \label{fig:no_coupling}
\end{figure}

We would like to assign the vibronic peaks from our simulation. To this end,
we run the simulation once again, this time, however we set the linear
diabatic constant to zero, i.e., $\lambda = 0$. With that change we can
reproduce the spectrum (at an equivalent level of theory) but in an artificial
case where there is no vibronic coupling. Figure~\ref{fig:no_coupling}
presents this spectrum. The non-coupled spectrum is easy to assign using the
labels that mark the symmetry of the electronic state, A$_1$ or B$_2$, and the
vibrational state ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of
quanta in mode $i$ with $i=1$ for the symmetric stretch, $i=2$ for the
symmetric bend and $i=3$ for the asymmetric stretch. The assigned spectrum
shows progressions in the symmetric bend. There is one such progression in
each electronic states.  Additionally for each state, there is also another
progression in the symmetric bend with one excitation in the symmetric
stretch.

\begin{figure}
    \includegraphics[width=16 cm]{./figures/sim_vs_zeke}
    \caption{
        Comparison of the simulated spectrum with the PFI-ZEKE
        experiment.\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
        shifted by $21$~meV$ = 170$~cm$^{-1}$ towards higher energies.
    }
    \label{fig:sim_vs_zeke}
\end{figure}

We decompose the vibronic states of our main simulation in the basis of the
artificial, uncoupled states discussed in the previous paragraph.
Figure~\ref{fig:sim_vs_zeke} shows the assigned spectrum and
Table~\ref{tab:peak_assignment} lists the decomposition of all peaks in the
region of up to $3000$~cm$^{-1}$ with intensities larger than $10^{-3}$. We
compare the assigned spectrum against the high-resolution
pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
2005.\cite{Willitsch:O3ZEKE:2005} The PFI-ZEKE spectrum is the best source of
information on the location of the peaks, especially the origin, against which
align our simulation. The origin of the simulated spectrum is located
$170$~cm$^{-1}$ lower than the experimental origin which was experimentally
observed at $101,020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
our error estimate for the vertical ionization energy
($30$~meV$=240$~cm$^{-1}$).  Once the origin locations are aligned the
comparison shows a good match between the simulation and the experiment.
Especially the states with an oscillator strength originating from the A$_1$
state are well aligned with the experimental features.

States close to the origin of the band are similar to the non-coupled states.
The origin peak is of a clear A$_1(0,0,0)$ character, the first two
excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very
similar to the non-coupled states, while the higher excitations in this
progression are showing large mixing. The same progression with one
vibrational quantum in the symmetric stretch is more interesting. The first of
its peaks is very well aligned with an experimentally visible feature, which
was previously assigned as the origin of the second state. The second peak in
this series, the first combination state A$_1(1,1,0)$, has very high
similarity to its uncoupled version. It is a good candidate to assign the
experimental, unassigned feature above $102,600$~cm$^{-1}$.

States with an oscillator strength from the B$_2$ state are on the other hand
not aligning well with the experiment. First such peak, close to the $12.64$~eV
mark on the Figure~\ref{fig:sim_vs_zeke}, is of a vibronic character and we
assign it as a A$_1$(0,0,1) state which gains intensity thanks to the coupling
to the B$_2$ state. This vibronic peak lies in the part of the spectrum marked
as pollution due to water. The origin of the B$_2$ state is very similar to
the uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
experimental spectrum. The next peak, slightly above the $12.76$~eV mark on
Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the symmetric
bend in the B$_2$ state, but as the label on the figure shows, it is only
about $55$\% similar to the uncoupled state.

\begin{figure}
    \includegraphics[width=16 cm]{figures/sim_vs_TarroniCarter}
    \caption{
        Comparison of the simulated spectrum to an earlier accurate simulation
        of Tarroni and Carter (assigned lines).~\cite{tarroni:O3:2011}
    }
    \label{fig:sim_vs_tarronicarter}
\end{figure}

Finally, we compare our simulation to an earlier accurate simulation, the work
of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011} Both spectra show a
good match to the experiment and agree in the assignment of the first
progression in the bending mode. The previous simulation, however, shows
smaller spacing between other lines, showing much higher congestion of the
spectrum. Additionally the origin of the second state falls at lower energy
aligning well with the peak that was assigned in the PFI-ZEKE also as the
origin of the second state.  While an additional simulation of the line
intensities would allow for the most complete comparison to the experimental
spectrum, the strong alignment of the simulated peaks of a leading
A$_1$ character with the experimental features leads us to believe that our
simulation offers the best assignment of the spectrum to date.

\begin{table}
    \renewcommand{\arraystretch}{0.8}% Tighter
    \caption{Assignment and decomposition of the eigenvectors of the ozone
    cation.}
    \label{tab:peak_assignment}
    \begin{tabular}{|l|l|l|}
        \hline
        Peak position (cm$^{-1}$) & Assignment & Eigenvector \\
        \hline
$     0$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
$   618$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
$   915$ &  $A_1(001)$ & \\
$  1076$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
$  1222$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
$  1368$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
$  1498$ &             & $ -0.10~B_2(000)$ \\
$  1684$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
$  1796$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
$  1848$ &             & $ +0.31~A_1(030)$ \\
$  1921$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
$  1977$ &             & $ +0.24~B_2(010)$ \\
$  2085$ &             & $ -0.52~B_2(010)$ \\
$  2132$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
$  2275$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
$  2363$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
$  2439$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
$  2454$ &             & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
$  2576$ &             & $ +0.27~B_2(020)$ \\
$  2736$ &             & $ +0.60~B_2(020)$ \\
$  2886$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
$  2974$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
$  3047$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
        \hline
    \end{tabular}
\end{table}

\section{Summary and Conclusions}

We have simulated the vibronic effects in the photoelectron spectrum of the
ozone.  We have used the KDC Hamiltonian\cite{Cederbaum:LVC:84, KDC:81,
Koppel:CIbookCh7:04} with the \emph{ab initio} parametrisation of the
quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
The results of our simulation match well the photoelectron spectrum from
1970s~\cite{dyke:O3:74} and the PFI-ZEKE spectrum from
2005.~\cite{Willitsch:O3ZEKE:2005} We present analysis of the spectra and
highlight the role of the vibronic coupling effects. Our simulation offers a
state-of-the-art insight into the spectrum and allows us to assign the
simulated peaks. Our assignment agrees well with the previously assigned
progression in the symmetric bend but we reassign some of the other features
of the spectrum. Simulated peaks that gain intensity from the B$_2$ state are
absent in the PFI-ZEKE spectrum which offers an interesting avenue for further
investigations. Modeling of the intensity of the PFI-ZEKE peaks is not
unlikely to bring insight into this problem.  Additionally our simulation does
not model dissociation of the ozone cation, which will play a role in the high
energy part of the spectrum.

\section{Acknowledgments}

This work was funded by the NSF Center for Chemical Innovation Phase I (grant
no. CHE-2221453) and U.S. Department of Energy, Basic Energy Sciences (grant
no. DE-FG02-05ER15629).

\clearpage

\bibliographystyle{prf}
\bibliography{allrefs}

\end{document}

% vim: tw=78 spell: