### 2024-09-02 article.tex
Update Resuts and Disucussion with Dr. Stanton comments

\documentclass[
12pt,
a4paper,
prb,
superscriptaddress,
% tightenlines,  % single-spaced
]{revtex4}
\usepackage{amsmath,amssymb,physics}
\usepackage{graphicx} % For figures with images
\usepackage{natbib} % bibliography
% \setcitestyle{super, sort&compress, comma} % chemistry style (prb's default)
\usepackage{hyperref} % Clickable links in the document and their colors
\hypersetup{
    colorlinks=true, % false: boxed links; true: colored links
    linkcolor=cyan, % color of internal links
    citecolor=magenta, % color of links to bibliography
    filecolor=magenta, % color of file links
    urlcolor=cyan, % color of external links
    runcolor=cyan
}

\begin{document}

\title{Vibronic Coupling Effects in the Photoelectron Spectrum of Ozone: A
Coupled-Cluster Approach}

\author{Pawe{\l} W{\'o}jcik}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\author{Anna I. Krylov}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{Hannah Reisler}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{P{\'e}ter G. Szalay}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\author{John F. Stanton}
\email[Electronic address: ]{jfstanton137@gmail.com}
\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\date{\today}

\begin{abstract}
One of the most important areas of application for equation-of-motion coupled
cluster (EOM-CC) theory is the prediction, simulation and analysis of
various types of electronic spectra.   In this work, the EOM-CC method known
as EOMIP-CC is applied to the closely lying and coupled pair of states of the
ozone cation -- ${\tilde X}^2A_1$ and ${\tilde A}^2B_2$ -- using sophisticated
approaches that extend to the full singles, doubles, triples and quadruples
model (EOMIP-CCSDTQ).   Combined with a venerable and powerful method for
calculating vibronic spectra from the Hamiltonian produced by EOMIP-CC
calculations, this research provides a spectrum that is in good agreement with
the photoelectron spectrum of ozone.   An important result here is that the
calculations suggest that the adiabatic gap separating these two electronic
states is somewhat smaller than currently thought; an assignment of the
simulated spectrum together with the more precise band positions of the
experimental measurements suggests that this energy gap is 1366$\pm$65
cm$^{-1}$.
\end{abstract}

\maketitle

\section{Introduction}

Amongst the brotherhood of triatomic molecules, it cannot be denied that water
(H$_2$O) is the most important, the most highly studied, and the most
well-understood.  Beyond H$_2$O, there are many triatomic molecules that have
an environmental, technological or biological importance, have been subjected
to many studied and are understood to various levels of detail.  Perhaps the
most interesting such case is ozone (O$_3$), which has a vast number of
important properties, a very rich history of study \cite{chappuis}, and 
unlike the relatively simple water molecule, a profound quantum-mechanical
complexity \cite{Babikov:anomalousOzone:2003}.  In the latter context, while
we think of (and an NMR experiment would reveal) as two distinct kinds of
oxygen atoms in ozone, the full molecular Hamiltonian does not distinguish
between them types, with the three equivalent structures separated by a
barrier that lies tantalizingly close to the O$_3$ $\rightarrow$ O$_2$ + O
dissociation threshold (102.4 kJ/mol) \cite{Ruscis:ATcT:2022}. In reality, the
energy levels of ozone all have a near triple ($e+a$) degeneracy, albeit with
a tunneling splitting so small that it can be ignored, along with a
semi-infinite lifetime (despite opposing opinion
\cite{Boggs:BerryOzone:2006}).   More than a half century ago, this intriguing
aspect of the ozone molecule was first discussed by Berry
\cite{Berry:Ozone:1960}.

Beyond the structural aspects of ozone, other mysteries center around this
curious molecule.  For example, the distribution of the eighteen distinct
$^{16}$O/$^{17}$O/$^{18}$O isotopologues in the Earth's atmosphere differs
from what is expected based on the natural isotopic abundance, a puzzle that
has been open for more than three decades
\cite{Mauersberger:OzoneMystery:1990}.

Among quantum chemists, ozone has a notorious history, with its strong
biradical character causing significant difficulties in calculations of its
ostensibly simple ground state molecular properties.  An early 1989 study
\cite{Stanton:Ozone:1989, Stanton:Ozone:1989b} by the Bartlett group and
collaborators found that the CCSD+T(CCSD) method predicted that the molecular
equilibrium structure of ozone would have $C_s$ symmetry (that is, the
asymmetric stretching harmonic frequency predicted by this method is
imaginary), a finding that led to a search for better treatment of
non-iterative triple excitations, ultimately leading to the well-known CCSD(T)
treatment \cite{Raghavachari:89, Urban:ccsd(t):1985, GenrefCCSD(T):93}. While
CCSD(T) and higher-level coupled-cluster methods available today
\cite{kucharski:ccsdtq:1992, Kallay:CCHigh, Matthews:ncc:2015} do a good job
in predicting the molecular vibrational potential, an elaborate multireference
configuration interaction study by Dawes {\it et al.} has done an excellent
job on the entire ground-state surface out to the dissociation limit
\cite{Dawes:ozone:2013}. As such, the quantum chemical understanding and
fidelity for the ground electronic state of O$_3$ is now at a mature level.

Qualitatively, the challenge posed to electronic structure theory by ozone
ultimately arises from its closely spaced highest-occupied ($a_2$) and
lowest-unoccupied ($b_1$) molecular orbitals (HOMO and LUMO, respectively).
The two electron configurations $[\cdots]b_2^2 a_1^0$ and $[\cdots]b_2^0
a_1^2$ mix strongly, posing the aforementioned challenges with (especially
single-reference) quantum chemical methods.   A second feature consequence of
the identity and energetic proximity of these two orbitals is that the ozone
cation (which is isoelectronic to the NO$_2$ radical) has closely lying
$^2A_1$ and $^2B_2$ electronic states.   Like the associated states in NO$_2$,
both of these states are plagued by orbital symmetry-breaking problems
\cite{Davidson:SymmBreak:76}, a problem that greatly complicates
quantum-chemical efforts to characterize them.   One of the many
accomplishments of the Bartlett group has been the integral role played by
them in the development of equation-of-motion coupled-cluster theory
\cite{Stanton:93:EOMCC, Nooijen:EOMEA:95, Bartlett:CC_review:07} (EOM-CC, also
known as linear response coupled-cluster theory \cite{Koch:90:LinResp}). These
methods provide a very efficient and simple way to study certain classes of
what are often termed ``multireference problems" \cite{Krylov:EOMRev:07}, and
are ideally suited to studying many reactive intermediates, radicals,
biradicals and electronically excited states.   From a somewhat wider
viewpoint, the existence of closely spaced electronic states always carries
the potential for (possibly strong) vibronic coupling, a phenomenon that can
play an important role in molecular dynamics and spectroscopy.  Indeed, one of
the great successes of EOM-CC methods has been in their ability -- if and only
if combined with vibronic coupling models -- to enable high-quality
simulations of complicated electronic spectra.   Such work provides important
insight into the nature of vibronic coupling in molecular systems, as has been
exemplified by various application studies (for example, see
Refs.~\cite{Stanton:NO3:07} and~\cite{Koppel:02}).

Our contribution to this issue paying homage to the career and accomplishments
of R.J. Bartlett consists of an application of a vibronic coupling model,
parametrized by EOM-CC calculations, to the photoelectron spectrum of ozone.
We trust that this combination of methodology applied to a molecule that has
been extensively studied by the Bartlett group is an appropriate contribution.

\begin{figure}
    \includegraphics[width = 16 cm]{./figures/ozone_intro}
    \caption{ 
        Two lowest states of the ozone cation and normal modes of the ozone
        neutral.
    }
    \label{fig:ozone_intro}
\end{figure}

\section{Theory}

Ozone is a $C_{2v}$ molecule (following Mulliken's
convention~\cite{Mulliken:55:symnot} the molecule is placed in the $yz$-plane
and the molecular symmetry axis aligns with the $z$ direction) with three
normal modes: symmetric stretch, symmetric bend, and asymmetric stretch.  The
asymmetric stretch is of $b_2$ symmetry. The two lowest electronic states of
the ozone cation are close in energy. The lower state is fully symmetric while
the higher one is of $B_2$ symmetry, the same symmetry as the asymmetric
stretch. The close energetic separation and the matching symmetry of the
asymmetric stretch is accompanied by significant vibronic effects in the
lowest two states of the ozone photoelectron.  Figure~\ref{fig:ozone_intro}
presents a graphical summary of this information.

The vibronic states of the ozone cation are simulated using the model
Hamiltonian of K{\"o}ppel, Domcke, and Cederbaum (KDC
Hamiltonian).\cite{Cederbaum:LVC:84,KDC:81,Koppel:CIbookCh7:04} This is a
multi-state and multi-mode Hamiltonian defined in the basis of diabatic
states. The ozone cation is considered in the basis of two quasi-diabatic
states coupled by one mode (mode number 3). The model also includes two
symmetric modes (modes number 1 and 2)
\begin{subequations}
    \begin{equation}
        H = H _0 \mathbf{1}
        +
        \begin{pmatrix}
            V ^{(1)}  & \lambda _3 Q _3\\
            \lambda _3 Q _3 & V ^{(2)}
        \end{pmatrix}
    \end{equation}
    \begin{equation}
        H _0 = 
        \frac{1}{2} \left(\sum _{i = 1}^3 
        - \omega _i \frac{\partial ^2}{\partial Q _i ^2 }\right)
        + \frac{1}{2}\omega _3 Q _3 ^2
    \end{equation}
    \begin{equation}
        V ^{(\alpha)} = 
        E ^{(\alpha)} 
        + \sum _{i \in \{1,2\}} 
            \kappa ^{(\alpha)} _i Q _i 
        + \sum _{i,j \in \{1,2\}} 
            \kappa ^{(\alpha)} _{ij} Q _i Q _j 
        + \sum _{i,j,k \in \{1,2\}} 
            \kappa ^{(\alpha)} _{ijk} Q _i Q _j Q _k 
        + \sum _{i,j,k,l \in \{1,2\}} 
            \kappa ^{(\alpha)} _{ijkl} Q_i Q _j Q _k Q _l.
    \end{equation}
    \label{eq:KDC}
\end{subequations}
The vibronic Hamiltonian elements are expanded around the equilibrium geometry
of neutral ozone. $E ^{(\alpha)}$ are the vertical ionization energies
calculated at that geometry. $Q_i$ are the dimensionless normal coordinates of
ozone, $\kappa$ are the coefficients of expansion of the potential along the
fully symmetric coordinates, and $\lambda$ the linear diabatic coupling,
$\omega$ represent the harmonic frequencies of the neutral.

The parameters that enter the KDC Hamiltonian are computed using \emph{ab
inito} coupled-cluster (CC) methods and its equation-of-motion EOM-CC
extension.~\cite{Bartlett:CC_review:07, Krylov:EOMRev:07, Bartlett:Book:09,
Christiansen:EOMRev:11, Bartlet:EOMRev:12, Krylov:OSRev} The CC expansion is
truncated to singles and doubles (CCSD), singles, doubles and triples (CCSDT)
as well as singles, doubles, triples and quadruples
(CCSDTQ).\cite{Matthews:ncc:2015}  The states of the cation are described with
EOM-CC for ionization potential (EOMIP).~\cite{StantonGauss:EOMIP:99} The
definition of quasi-diabatic states of Ichino, Gauss and Stanton based on the
EOMIP-CC method (EOMIP-CC-QD) is used here~\cite{Stanton:EOMIPdeg:09}. All CC
and EOM-CC calculations correlate all electrons, in other words, the
frozen-core approximation is not invoked.

The geometry of the neutral ozone and the computation of its harmonic
frequencies and normal coordinates uses
CCSDT/ANO1.~\cite{Almlof:ANO,Almlof:ANO:1988} 
The linear diabatic coupling $\lambda$ is computed at the same geometry using
EOMIP-CCSD-QD/ANO1. The expansion coefficients $\kappa$ are computed using the
EOMIP-CCSD/ANO1 energies calculated on a grid around the same geometry.

The vertical ionization energies are calculated using a composite method.  The
starting value is the complete basis set (CBS) extrapolation of the
EOM-IP-CCSDT/cc-pCVnZ energies with n = 5, 6.~\cite{Woon:95:CCBS} These values
are augmented with two corrections: the $\Delta$Q correction in the cc-pwCVTZ
basis set and the relativistic correction calculated using
EOM-IP-CCSD/cc-pwCVTZ.~\cite{Dunning:02:p(w)CVnZ} The error estimate of the
extrapolated CBS reference values is reported using half of the absolute value
of the difference between the best \emph{ab initio} value and the extrapolated
value.  For the remaining corrections (quadruples correction, $\Delta Q$, and
the relativistic correction) the error estimate is reported as half of the
absolute value of the correction.

To compare the simulated photoelectron spectrum to experiment, the
cross-section ratio A$_1$:B$_2$ of $1$:$1.35$ is used.\cite{KDC:O3:92}
Additionally, the stick spectrum is broadened with the Lorenzian envelopes
normalized to the feature intensities
\begin{equation}
    f _{envelope}(x, x _i, I _i) = 
    \frac{I_i}{(x-x _i)^2 + (\gamma/ 2) ^2}.
    \label{eq:lorentzian}
\end{equation}
$x _i$ is the position of the spectral peak, $I _i$ is its intensity and
$\gamma$ is the peak's width.

The spectrum is simulated using the \textsc{xsim} program in \textsc{CFOUR},
with the basis of 50 harmonic states in each mode and 6000 iterations of the
Lanczos procedure.~\cite{Sharma:xsim_socjt:2024} All calculations are
completed using \textsc{CFOUR}.~\cite{cfour, cfour:2020}

\section{Results}

The optimized geometry of ozone gives a bond length of $1.270$~\AA{} and a
bond angle of $116.9^\circ$. The two symmetric normal modes have frequencies
$1169$~cm$^{-1}$ and $724$~cm$^{-1}$, while the asymmetric stretch has a
frequency equal to $1129$~cm$^{-1}$. The computed value of the linear diabatic
coupling constant $\lambda$ is $1394$~cm$^{-1}$. The vertical ionization
energy for the first excited state, $E^{(1)}$, computed using the composite
method described above is  $12.827$~eV. Our error estimate for this value is
$30$~meV, see Table~\ref{tab:vertical_ionization_energy} for details.  We note
that the convergence of the vertical gap between the two states is much faster
and this value is estimated as $123\pm8$~meV, see Table~\ref{tab:vertical_gap}
for details.

\begin{table}
    \caption{
        Vertical ionization energies with the error estimates, eV.
    }
    \label{tab:vertical_ionization_energy}
    \begin{center}
        \begin{tabular}[c]{|l|rr|r|}
            \hline
            Contribution  & $^2$A$_1$ & $^2$B$_2$ & Error estimate \\ \hline
            CBS           & 12.872    & 12.981    & 0.02 \\
            $\Delta$Q     & -0.037    & -0.021    & 0.02 \\
            Relativistic  & -0.008    & -0.010    & 0.005 \\ \hline
            Sum           & 12.827    & 12.950    & 0.03 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{table}
    \caption{
        The energy gap between vertical ionization energies of the $^2$A$_1$
        and $^2$B$_2$ states with error estimates, meV. See caption of
        Table~\ref{tab:vertical_ionization_energy}.
    }
    \label{tab:vertical_gap}
    \begin{center}
        \begin{tabular}[c]{|l|r|r|}
            \hline
            Contribution             &  Vertical gap    & Error estimate \\ \hline
            EOM-CCSDT/CBS            &  108.8           & 0.9 \\
            $\Delta$Q/pwCVTZ         &   16.5           & 8 \\
            Relativistic/CCSD/pwCVTZ &   -2.2           & 1.1 \\ \hline
            Final value, meV         &  123             & 8 \\
            Final value, cm$^{-1}$   &  990             & 65 \\ \hline
        \end{tabular}
    \end{center}
\end{table}

\begin{figure}
\includegraphics[width = 8 cm]{figures/sim_vs_Dyke}
\caption{
    Comparison of the experimental (solid black line) and the simulated
    spectrum. In the simulated spectra the plot is generated with peak width
    $\gamma = 30$~meV. The simulated spectrum is shifted towards higher
    energies by $21$~meV.
}
\label{fig:sim_vs_dyke}
\end{figure}

Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
experimental one taken from reference~\cite{dyke:O3:74}. Our simulation allows
for an additional element of analysis of the simulated spectrum.
Figure~\ref{fig:ozone_overlay} presents a decomposition of the simulated
spectrum from Figure~\ref{fig:sim_vs_dyke}. All lines that contribute to the
spectrum are marked individually and are color-coded indicating which
electronic state's transition intensity the peak draws from. The total
envelope of the spectrum is also decomposed showing contributions from both
states. Our simulation locates the minimum of the conical intersection at
$3174$~cm$^{-1}$ above the origin ($12.92$~eV) that is marked on the figure as
CI.

\begin{figure}
    \includegraphics[width=10 cm]{figures/spectrum_overline}
    \caption{
        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
        scale in eV. Top axis shows energy offset from the origin in
        cm$^{-1}$. Stick spectrum shows positions and intensities of all
        simulated states. Colors blue and orange mark states of A$_1$ and
        B$_2$ symmetry, respectively. Gray vertical lines with captions on top
        indicate positions of features as measured by the PFI-ZEKE
        experiment.~\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
        shifted to match the PFI-ZEKE experimental origin. $D_0$ marks the
        dissociation threshold of O$_3^+$.~\cite{Willitsch:O3ZEKE:2005}
        Gray dotted line marks the energy of the minimum of the conical
        intersection (CI) located by our Hamiltonian.
    }
    \label{fig:ozone_overlay}
\end{figure}

\begin{figure}
    \includegraphics[width=10cm]{figures/spectrum_assigned.pdf}
    \caption{
        Simulated Franck-Condon photoelectron spectrum of ozone, vibronic
        coupling removed.
    }
    \label{fig:no_coupling}
\end{figure}

The vibronic peaks are assigned using a comparison to the Franc-Condon
simulation.  To this end, the spectrum is simulated once again, this time,
however, setting the linear diabatic constant to zero, i.e., $\lambda = 0$.
That change reproduces the spectrum (at an equivalent level of theory) but in
an artificial case where there is no vibronic coupling, i.e., the
Franck-Condon spectrum.  Figure~\ref{fig:no_coupling} presents this spectrum.
The non-coupled spectrum is easy to assign using the labels that mark the
symmetry of the electronic state, A$_1$ or B$_2$, and the vibrational state
($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of quanta in mode $i$
with $i=1$ for the symmetric stretch, $i=2$ for the symmetric bend and $i=3$
for the asymmetric stretch.  The assigned spectrum shows progressions in the
symmetric bend. There is one such progression in each electronic state.
Additionally for each state, there is also another progression in the
symmetric bend with one excitation in the symmetric stretch.

\begin{figure}
    \includegraphics[width=16 cm]{./figures/sim_vs_zeke}
    \caption{
        Comparison of the simulated spectrum with the PFI-ZEKE
        experiment.\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
        shifted by $21$~meV$ = 170$~cm$^{-1}$ towards higher energies.
    }
    \label{fig:sim_vs_zeke}
\end{figure}

The vibronic states are decomposed in the basis of the artificial, uncoupled
states discussed in the previous paragraph.  Figure~\ref{fig:sim_vs_zeke}
shows the assigned spectrum and Table~\ref{tab:peak_assignment} lists the
decomposition of all peaks in the region of up to $3000$~cm$^{-1}$ with
intensities larger than $1/300$ of the maximum peak. 
The resulting assignment is compared against the high-resolution
pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
Ref.~\cite{Willitsch:O3ZEKE:2005}. The PFI-ZEKE spectrum is the best source of
information on the location of the peaks, especially the origin, against which
the origin peak of the simulation is aligned. The origin of the simulated
spectrum is located $170$~cm$^{-1}$ lower than the experimental origin which
was seen at $101 020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
our error estimate for the vertical ionization energy
($30$~meV$=240$~cm$^{-1}$).

\begin{figure}
    \includegraphics[width=16 cm]{figures/sim_vs_TarroniCarter}
    \caption{
        Comparison of the simulated spectrum to an earlier accurate simulation
        of Tarroni and Carter (assigned lines).~\cite{tarroni:O3:2011}
    }
    \label{fig:sim_vs_tarronicarter}
\end{figure}

Lastly, a comparison to an earlier accurate simulation, the work of Tarroni
and Carter, Ref.~\cite{Willitsch:O3ZEKE:2005}, is presented on
Figure~\ref{fig:sim_vs_tarronicarter}. The comparison includes only the lines
from the earlier simulation that were tabulated with an assignment.

\begin{table}
    \renewcommand{\arraystretch}{0.8}% Tighter
    \caption{Assignment and decomposition of the eigenvectors of the ozone
    cation.}
    \label{tab:peak_assignment}
    \begin{tabular}{|l|l|l|l|}
        \hline
        Energy (cm$^{-1}$) & Symmetry& Assignment & Eigenvector \\
        \hline
$     0$ & $A_1$ &  $A_1(000)$ & $ -0.99~A_1(000)$ \\
$   618$ & $A_1$ &  $A_1(010)$ & $ +0.99~A_1(010)$ \\
$   915$ & $B_2$ &  $A_1(001)$ & \\
$  1076$ & $A_1$ &  $A_1(100)$ & $ -0.99~A_1(100)$ \\
$  1222$ & $A_1$ &  $A_1(020)$ & $ -0.97~A_1(020)$ \\
$  1368$ & $B_2$ &  $B_2(000)$ & $ +0.97~B_2(000)$ \\
$  1498$ & $B_2$ &             & $ -0.10~B_2(000)$ \\
$  1684$ & $A_1$ &  $A_1(110)$ & $ -0.96~A_1(110)$ \\
$  1796$ & $A_1$ &  $A_1(030)$ & $ -0.87~A_1(030) -0.16~A_1(110) -0.13~A_1(020) +0.13~A_1(040)$ \\
$  1848$ & $A_1$ &             & $ +0.31~A_1(030)$ \\
$  1921$ & $B_2$ &  $B_2(010)$ & $ -0.74~B_2(010) +0.18~B_2(020) +0.11~B_2(000)$ \\
$  1977$ & $B_2$ &             & $ +0.24~B_2(010)$ \\
$  2085$ & $B_2$ &             & $ -0.52~B_2(010)$ \\
$  2132$ & $A_1$ &             & $ +0.25~A_1(030) +0.24~A_1(040) +0.13~A_1(120) -0.13~A_1(050)$ \\
$  2143$ & $A_1$ &             & \\ 
$  2275$ & $A_1$ &  $A_1(120)$ & $ +0.86~A_1(120) +0.12~A_1(040)$ \\
$  2363$ & $A_1$ &             & $ -0.62~A_1(040) +0.40~A_1(120) -0.17~A_1(030) +0.11~A_1(050)$ \\
$  2436$ & $A_1$ &             & $ +0.47~A_1(040) +0.12~A_1(120)$ \\
$  2453$ & $B_2$ &  $B_2(020)$ & $ +0.53~B_2(020) +0.24~B_2(010) -0.17~B_2(030)$ \\
$  2576$ & $B_2$ &             & $ +0.27~B_2(020)$ \\
$  2735$ & $B_2$ &             & $ +0.60~B_2(020)$ \\
$  2743$ & $A_1$ &             & \\ 
$  2862$ & $A_1$ &             & \\ 
$  2903$ & $A_1$ &  $A_1(130)$ & $ -0.81~A_1(130) +0.12~A_1(050) -0.10~A_1(120)$ \\
$  2969$ & $A_1$ &             & \\ 
$  2977$ & $B_2$ &             & $ -0.24~B_2(110) -0.16~B_2(020)$ \\
$  3006$ & $B_2$ &             & \\ 
$  3045$ & $A_1$ &  $A_1(050)$ & $ +0.76~A_1(050) +0.13~A_1(130)$ \\
        \hline
    \end{tabular}
\end{table}

\section{Discussion}

We first discuss the comparison of our simulations to the lower-resolution
experimental spectrum presented on Figure~\ref{fig:sim_vs_dyke}. The peak A is
known to be a hot band, and does not appear in our ($0$~K)
simulation.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
increase in the intensities of peaks B, C, D, and E. The spacing between these
peaks is also well reproduced. The drop in the intensity at the peak F is
captured by the simulation. Starting from the peak G, the simulation shows
discrepancy with the experiment. A sudden drop in the intensity past the peak
H is not observed in the simulation. A likely source of this mismatch is
discussed later.

The decomposition of the spectrum presented on Figure~\ref{fig:ozone_overlay}
introduces additional insight. The spectrum shows that peaks B and C are
almost purely of the $^2$A$_1$ electronic and $a_1$ vibrational character.
Starting from the peak D, the contributions from two states are equally
important. It is also clear that the following peaks arise as mixtures of many
vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin the
density of vibronic peaks increases significantly. This value can be compared
to the minimum of the conical intersection located about $1000$~cm$^{-1}$
higher. Our simulation is incapable of accounting for dissociation of the
molecule, therefore we expect a discrepancy with the experiment as the energy
gets closer to the dissociation threshold located at $4898\pm3$~cm$^{-1}$
above the origin.~\cite{Willitsch:O3ZEKE:2005}

Moving on to the comparison with the high-resolution spectrum, see
Figure~\ref{fig:sim_vs_zeke}, reveals more details. Lines of the PFI-ZEKE
experiment and our simulation show a good match. Especially the states of the
A$_1$ symmetry are well aligned with the experimental features. States close
to the origin are similar to the non-coupled states. The origin peak is of
A$_1(0,0,0)$ character, the first two excitations in the symmetric bend,
A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very similar to the non-coupled states,
while the higher excitations in this progression are showing large mixing. The
same progression with one vibrational quanta in the symmetric stretch is more
interesting. The first of its peaks is very well aligned with experimentally
visible feature, which was previously assigned as the origin of the second
electronic state. The second peak in this series, A$_1(1,1,0)$, has very high
similarity to its uncoupled version. It is a good candidate to assign the
experimental, unassigned feature above $102,600$~cm$^{-1}$.

States of the B$_2$ symmetry are, on the other hand, not aligning well with
the experiment. The first such peak, close to the $12.64$~eV mark on
Figure~\ref{fig:sim_vs_zeke}, is assigned to the A$_1$(0,0,1) state.  This is
a vibronic state that gains intensity thanks to the coupling to the electronic
$^2$B$_2$ state. This vibronic peak lies in the part of the spectrum where
bands of oxygen (which appears as impurity in the gas sample) are visible in
the experiment. The origin of the $^2$B$_2$ state is very similar to the
uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
experimental spectrum. The next peak, slightly above a $12.76$~eV mark on
Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the
symmetric bend in the $^2$B$_2$ state, but as the label on the figure shows,
it is only about $55$\% similar to the uncoupled state.

The comparison to the earlier simulation on
Figure~\ref{fig:sim_vs_tarronicarter} shows that both present a good match to
the experiment. Both simulations also agree in the assignment of the first
progression in the bending mode. The previous simulation, however, shows
smaller spacing between the remaining lines, showing much higher congestion of
the spectrum. Additionally the origin of the second state falls at lower
energy aligning well with the peak that was assigned in the PFI-ZEKE
experiment also as the origin of the second state. While an additional
simulation of the line intensities would allow for the most complete
comparison to the experimental spectrum, the strong alignment of the simulated
peaks of a A$_1$ symmetry with the experimental features leads us to
believe that our simulation offers the best assignment of the spectrum to
date.

\section{Summary and Conclusions}

In this exploration of the photoelectron spectrum of the ozone molecule,
cutting-edge high-order coupled cluster models have been used together with a
vibronic Hamiltonian approach to spectroscopy beyond the Born-Oppenheimer
approximation \cite{Cederbaum:LVC:84, KDC:81, Koppel:CIbookCh7:04}.
Calculations based on the equation-of-motion method known as EOMIP-CC have
been used to parametrize the Hamiltonian, which relies importantly on the
quasi-diabatic states of Ichino, Gauss and Stanton.~\cite{Stanton:EOMIPdeg:09}
The results of the simulation provide an excellent match to the experimental
spectrum from the ionization potential of the lower state to about 2500
cm$^{-1}$ higher (the restriction arising from the local parametrization of
the Hamiltonian).  Analysis of the spectrum and its simulation underscore the
importance of vibronic coupling effects in the ozone cation.  Results of this
work offer a state-of-the-art insight into the details of the spectrum and
allows for an assignment of the experimental results.  Simulated peaks that
gain intensity from the ${\tilde A}B_2$ state are absent in the PFI-ZEKE
spectrum which offers an interesting avenue for further investigations.

\section{Acknowledgments}

This project was initiated when three of the authors (P.W, P.G.S. and J.F.S)
were in Budapest, where J.F.S. was serving as a recipient of the John von
Neumann Award in STEM bestowed by the Fulbright Foundation. Additional
research presented here benefited from the NSF Center for Chemical Innovation
Phase I (grant no. CHE-2221453) and the U.S. Department of Energy, Basic Energy
Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
Prof. Bartlett a happy ninth decade of life, and hope that he continues to
stimulate others in the field with his creative insights.

\clearpage

\bibliography{allrefs}

\end{document}

% vim: tw=78 spell: