### 2024-09-04 si.tex

\documentclass[12pt,prb,aps]{revtex4-1}
\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{verbatim}

\usepackage{xr}
\externaldocument[si-]{si}

\renewcommand\thesection{\arabic{section}}
\renewcommand\thesubsection{\thesection.\arabic{subsection}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thepage}{S\arabic{page}}

%Number the pages consecutively using the format S1, S2, etc..
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{\thepage}

\begin{document}

\title{Vibronic Coupling Effects in the Photoelectron Spectrum of Ozone: A
Coupled-Cluster Approach:\\
Supporting Information}

\author{Pawel Wojcik}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}
%AIK: Not sure this is appropriate. If you just visited for a few weeks, does not make an affiliation.
%\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A, Budapest, Hungary}

\author{Hanna Reisler}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{P{\'e}ter G. Szalay}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\author{Anna I. Krylov}
\affiliation{Department of Chemistry, University of Southern California, Los Angeles, California 90089, USA}

\author{John F. Stanton}
\email[Electronic address: ]{jfstanton137@gmail.com}
\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

\maketitle

\renewcommand{\baselinestretch}{1.0}  %for tighter appearance
\section{xsim input}
Full set of parameters used to define the KDC Hamiltonian. The parameters are presented as an input file to the \textsc{xsim} program.
\begin{verbatim}
Units
1
Dataset
1
States
2
Modes
3
Basis Functions
50 50 50
Lanczos
6000
Transition Moment
1. 1.35 ! Muller, Koppel, Cederbaum, et al. Chem. Phys. Lett. 197, 599 (1992)
Vertical Energies, cm-1
1 103456.7 ! EOMIP-CCSDT/pCVnZ/CBS n=5,6 + dQ/pwCVTZ + relativistic/CCSD/pwCVTZ
2 104449.7 ! EOMIP-CCSDT/pCVnZ/CBS n=5,6 + dQ/pwCVTZ + relativistic/CCSD/pwCVTZ
0 0 0 0 0 0 0 0 0 0 0 0
Harmonic Frequencies, cm-1 ! CCSDT/ANO1 for the neutral state
1 724.19
2 1168.58
3 1128.50
0 0 0 0 0 0 0 0 0 0 0 0
Linear, cm-1 ! Everything below: EOM-CCSDT/ANO1
1 1 2     1178.18
1 1 1     1261.02
2 2 2     -781.75
2 2 1    -1303.82
1 2 3     1393.82
0 0 0 0 0 0 0 0 0 0 0 0
Quadratic, cm-1
1 1 2 2      802.90
1 1 1 2     -185.50
1 1 1 1      779.81
2 2 2 2     1276.61
2 2 1 2       23.56
2 2 1 1      369.33
0 0 0 0 0 0 0 0 0 0 0 0
Cubic, cm-1
1 1 2 2 2     -202.85
1 1 1 2 2       74.39
1 1 1 1 2      -62.49
1 1 1 1 1       90.19
2 2 2 2 2     -282.04
2 2 1 2 2       75.56
2 2 1 1 2       -0.49
2 2 1 1 1       71.32
0 0 0 0 0 0 0 0 0 0 0 0
Quartic, cm-1
1 1 2 2 2 2       34.17
1 1 1 2 2 2      -26.65
1 1 1 1 2 2        3.16
1 1 1 1 1 2      -12.25
1 1 1 1 1 1        3.92
2 2 2 2 2 2       41.58
2 2 1 2 2 2      -35.55
2 2 1 1 2 2       -7.14
2 2 1 1 1 2       -6.88
2 2 1 1 1 1       11.34
0 0 0 0 0 0 0 0 0 0 0 0
\end{verbatim}

\end{document}